Perceptual Symbol Grounding in \\Holographic Memories

### 2016-12-16 papers/hrr-paper/paper.tex
Minor correction to text and image positioning and sizes.

	\begin{figure}
		\center
		\includegraphics[width=0.9\columnwidth]{img/scalar-probe.png}
		\caption{Result of $(80 \otimes A) \oslash A$ for a 2000-element HRR encoding values from range [0,100].}
	\end{figure}

@@

	- Implementation and building up of HRRs (similarity measures and cosine, lookup table...)

@@

	- necessity (benefit of having it, querying smth that was not there)
	- implementation

@@

	Encoding scalars is a good example for the necessity of preserving topology.
Depending on what dataset is in use it may be required to store mappings for any number of values, which can drastically increase space requirements.
Furthermore, using a separate random encoding for each scalar, regardless of whether they are very close to each other or not, there would be no similarity in neighbouring values.
As such the following probe will yield no result:

@@

	\begin{figure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-pre-perm.png}
			\caption{Binding}
			\label{no-perm-a}
		\end{subfigure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-pre-perm-probe.png}
			\caption{Probing}
			\label{no-perm-b}
		\end{subfigure}
		\caption{Plot of vector values.
Figure \ref{no-perm-a} shows the binding of the value 8 on a scale to 10 with the symbolic representation of "ball", which is random.
In figure \ref{no-perm-b} "ball" is being probed, and the expected result is a Gaussian distribution for 8, mixed with some noise.}
		\label{no-perm}
	\end{figure}

@@

	\begin{figure}
		\includegraphics[width=\columnwidth]{img/scalar-pre-perm-fft.png}
		\includegraphics[width=\columnwidth]{img/scalar-random-fft.png}
		\caption{Fourier analysis of 1D symbols: Gaussian distribution (top row) and same dimensional vector randomly sampled from normal distribution (bottom row). It is obvious that latter is uniformly spread across the whole vector, while the former is concentrated around 2 boundary values leading to increased interference between multiple items stored one over another (when added in Fourier space).}
		\label{fft}
	\end{figure}

@@

	\begin{algorithm}
		\caption{Original pipeline}
		\begin{algorithmic}[1]
			\ForAll{$symbol$}
			\State $t_{symbol} \gets chooseRandomVector()$
			\EndFor
			\State $m \gets initEmptyMemory()$
			\ForAll{$(symbol_1, symbol_2)$}
			\State $enc_1 \gets FFT(t_{symbol_1})$
			\State $enc_2 \gets FFT(t_{symbol_2})$
			\State $m \gets m + IFFT(enc_1 \otimes enc_2$)
			\EndFor		
			\Function{ProbeMemory}{memory, operand}
			\State $result\gets IFFT(FFT(memory)\oslash FFT(operand))$
			\ForAll{$symbol$}
			\State $s \gets cosineSimilarity(t_{symbol}, result)$
			\If {$s \geq cleanResultSimilarity$}
			\State $cleanResultSimilarity \gets s$
			\State $cleanResult \gets t_{symbol}$
			\EndIf
			\EndFor \\
			\Return $cleanResult$
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	Lines 1-3 assign a random vector representation to each symbol known to the system prior to any other operations. This lookup table is stored in array $t$. Lines 4-9 initialize the memory $m$ with arbitrary amount of pairs of associated scalars. Function $ProbeMemory(memory, operand)$ returns symbol associated with provided $operand$. Line 11 performs circular correlation and stores noisy result of unbinding in $s$. Lines 12-18 loop through the whole lookup table and find the maximum similarity vector to $s$.

@@

	\begin{algorithm}
		\caption{Modified pipeline}
		\begin{algorithmic}[1]
			\State $P \gets chooseRandomPermutation(N)$
			\State $m \gets initEmptyMemory()$
			\ForAll{$(symbol_1, symbol_2)$}
			\State $enc_1 \gets FFT(permute(enc(symbol_1, N), P))$
			\State $enc_2 \gets FFT(permute(enc(symbol_2, N), P))$
			\State $m \gets m + enc_1 \otimes enc_2$
			\EndFor	
			\Function{ProbeMemory}{memory, operand}
			\State $r \gets permute(IFFT(memory \oslash operand), P^{-1})$
			\State $cleanResult \gets argmax(r)$

@@

@@

	Just as Plate has done in the original HRR paper \cite{Plate:1995:HolographicReducedRepresentations} we have tested the memory capacity in terms of accuracy of successful recall for variable numbers of bound pairs. Results are shown in figure \ref{fig:capacity}.
		\begin{figure}[th!]
		\begin{subfigure}{\columnwidth}
			\center
			\includegraphics[width=0.8\columnwidth]{img/capacity.png}
			\caption{Memory capacity analysis for storing a variable number of pair associations in a single memory vector. Different lines show various vector lengths, the horizontal axis shows the number of pairs and vertical axis accuracy of successful recall. Both false positives and false negatives are regarded.}
			\label{fig:capacity}
		\end{subfigure}
				\begin{subfigure}{\columnwidth}
			\center
			\includegraphics[width=0.8\columnwidth]{img/capacity_scalar_30.png}
			\caption{Memory capacity analysis for binding pairs of one scalar and a random symbol, then probing for the scalar.}
			\label{fig:capacity_scalar}
		\end{subfigure}	
		\begin{subfigure}{\columnwidth}
			\center
			\includegraphics[width=0.8\columnwidth]{img/capacity_coordinate.png}
			\caption{Memory capacity analysis for binding pairs of one 2D coordinate and a random symbol, then probing for the scalar.}
			\label{fig:capacity_coordinate}
		\end{subfigure}	
		\end{figure}	
	As we can see, a 512 element vector starts decreasing accuracy of recall after 10 pairs stored in a single memory block. A 4096 and 8192 element vectors can almost perfectly recall 20 associated pairs, which matches the results of Plate using a lookup table and a clean-up process.

@@

	\begin{figure}[th!]
		\center
		\begin{subfigure}{0.8\columnwidth}
			\includegraphics[width=\columnwidth]{img/map_original.jpg}
			\caption{Original scene and set of detected objects}
			\label{fig:object-detection}
		\end{subfigure}

		\begin{subfigure}{0.45\columnwidth}
			\center
			\includegraphics[width=\linewidth]{img/coord_example_1.png}
			\caption{Encoding of (x,y) location for "watch"}
			\label{fig:encoding-coord}
		\end{subfigure}
		\begin{subfigure}{0.45\columnwidth}
			\center
			\includegraphics[width=\linewidth]{img/coord_example_2.png}
			\caption{Encoding of "watch" object ID as randomly sampled vector}
			\label{fig:encoding-object}
		\end{subfigure}

				\begin{subfigure}{0.8\columnwidth}
					\center
					\includegraphics[width=\linewidth]{img/coord_example_3.png}
					\caption{Representation of the complete visual scene, containing information about all 7 objects and their locations scrambled together}
					\label{fig:encoding-scene}
				\end{subfigure}

@@

    			\caption{Test}
    \end{figure}

@@

	% use section* for acknowledgment
	\section*{Acknowledgment}

	% leaving this as is.
	The authors would like to thank...
Manchester developers. :slowclap:

>>>>

@@

@@

@@

	Encoding scalars is a good example for the necessity of preserving topology.
Depending on what dataset is in use it may be required to store mappings for any number of values, which can drastically increase space requirements.
Furthermore, using a separate random encoding for each scalar, regardless of whether they are very close to each other or not, there would be no similarity in neighbouring values.
For example, the following probe will yield no result:

@@

	\begin{figure}[th!]
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-pre-perm.png}
			\caption{Binding}
			\label{no-perm-a}
		\end{subfigure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-pre-perm-probe.png}
			\caption{Probing}
			\label{no-perm-b}
		\end{subfigure}
		\caption{Plot of vector values.
Figure \ref{no-perm-a} shows the binding of the value 8 on a scale to 10 with the symbolic representation of "ball", which is random.
In figure \ref{no-perm-b} "ball" is being probed, and the expected result is a Gaussian distribution for 8, mixed with some noise.}
		\label{no-perm}
	\end{figure}

@@

\begin{figure}[th!]
	\includegraphics[width=\columnwidth]{img/scalar-pre-perm-fft.png}
	\includegraphics[width=\columnwidth]{img/scalar-random-fft.png}
	\caption{Fourier analysis of 1D symbols: Gaussian distribution (top row) and same dimensional vector randomly sampled from normal distribution (bottom row). It is obvious that latter is uniformly spread across the whole vector, while the former is concentrated around 2 boundary values leading to increased interference between multiple items stored one over another (when added in Fourier space).}
	\label{fft}
\end{figure}

@@

	\begin{algorithm}
		\caption{Original pipeline}
		\begin{algorithmic}[1]
			\ForAll{$symbol$}
			\State $t_{symbol} \gets chooseRandomVector()$
			\EndFor
			\State $m \gets initEmptyMemory()$
			\ForAll{$(symbol_1, symbol_2)$}
			\State $enc_1 \gets FFT(t_{symbol_1})$
			\State $enc_2 \gets FFT(t_{symbol_2})$
			\State $m \gets m + IFFT(enc_1 \otimes enc_2$)
			\EndFor		
			\Function{ProbeMemory}{memory, op}
			\State $result\gets IFFT(FFT(memory)\oslash FFT(op))$

			\ForAll{$symbol$}
			\State $s \gets cosineSimilarity(t_{symbol}, result)$
			\If {$s \geq cleanResultSimilarity$}
			\State $cleanResultSimilarity \gets s$
			\State $cleanResult \gets t_{symbol}$
			\EndIf
			\EndFor \\
			\Return $cleanResult$
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

	Lines 1-3 assign a random vector representation to each symbol known to the system prior to any other operations. This lookup table is stored in array $t$. Lines 4-9 initialize the memory $m$ with arbitrary amount of pairs of associated scalars. Function $ProbeMemory(memory, op)$ returns symbol associated with provided $op$. Line 11 performs circular correlation and stores noisy result of unbinding in $s$. Lines 12-18 loop through the whole lookup table and find the maximum similarity vector to $s$.

		\begin{algorithm}
			\caption{Modified pipeline}
			\begin{algorithmic}[1]
				\State $P \gets chooseRandomPermutation(N)$
				\State $m \gets initEmptyMemory()$
				\ForAll{$(symbol_1, symbol_2)$}
				\State $enc_1 \gets FFT(permute(enc(symbol_1, N), P))$
				\State $enc_2 \gets FFT(permute(enc(symbol_2, N), P))$
				\State $m \gets m + enc_1 \otimes enc_2$
				\EndFor	
				\Function{ProbeMemory}{memory, op}
				\State $r \gets permute(IFFT(memory \oslash op), P^{-1})$
				\State $cleanResult \gets argmax(r)$

@@

@@

		\begin{figure}[th!]
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/capacity.png}
				\caption{Binding of pairs of random vectors (original pipeline).}
				\label{fig:capacity}
			\end{subfigure}
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/capacity_scalar_30.png}
				\caption{Binding pairs of one scalar and a random symbol, then probing for the scalar.}
				\label{fig:capacity_scalar}
			\end{subfigure}	
			\center
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/capacity_coordinate.png}
				\caption{Binding pairs of one 2D coordinate and a random symbol, then probing for the scalar.}
				\label{fig:capacity_coordinate}
			\end{subfigure}	
			\caption{Memory capacity analysis for storing a variable number of pair associations in a single memory vector. Different lines show various vector lengths, the horizontal axis shows the number of pairs and vertical axis accuracy of successful recall. Both false positives and false negatives are regarded.}
		\end{figure}

@@

	Just as Plate has done in the original HRR paper \cite{Plate:1995:HolographicReducedRepresentations} we have tested the memory capacity in terms of accuracy of successful recall for variable numbers of bound pairs. Results are shown in figure \ref{fig:capacity}.

	As we can see, a 512 element vector starts decreasing accuracy of recall after 10 pairs stored in a single memory block. A 4096 and 8192 element vectors can almost perfectly recall 20 associated pairs, which matches the results of Plate using a lookup table and a clean-up process.

@@

	\begin{figure}[th!]
		\center
		\begin{subfigure}{1.0\columnwidth}
			\center
			\includegraphics[width=0.6\columnwidth]{img/map_original.jpg}
			\caption{Original scene and set of detected objects}
			\label{fig:object-detection}
		\end{subfigure}

		\begin{subfigure}{0.45\columnwidth}
			\center
			\includegraphics[width=\linewidth]{img/coord_example_1.png}
			\caption{Encoding of (x,y) location for "watch"}
			\label{fig:encoding-coord}
		\end{subfigure}
		\begin{subfigure}{0.45\columnwidth}
			\center
			\includegraphics[width=\linewidth]{img/coord_example_2.png}
			\caption{Encoding of "watch" object ID as randomly sampled vector}
			\label{fig:encoding-object}
		\end{subfigure}
				\begin{subfigure}{0.8\columnwidth}
					\center
					\includegraphics[width=0.8\linewidth]{img/coord_example_3.png}
					\caption{Representation of the complete visual scene, containing information about all 7 objects and their locations scrambled together}
					\label{fig:encoding-scene}
				\end{subfigure}

@@

    			\caption{Similarity test for each symbol from the lookup table after probing the entire scene memory for a specific location.}
    \end{figure}

@@
